---
title: "Health Disparities"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(foreign)
library(knitr)
library(tidyr)
library(broom)
```

# Uploading the Demographic Data

```{r}
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/DEMO_J.XPT", tf <- tempfile(), mode="wb")

# Create Data Frame From Temporary File
demo <- foreign::read.xport(tf)
    
# save as an R data frame
# TutorialUser: update the file path here to a directory where you want to save the data frame 
saveRDS(demo, file="C:\\NHANES\\DATA\\DEMO_J.rds")  
```
## Removing Extraneous Demographic Variables

```{r}
demo <- demo %>%
  select(SEQN, RIAGENDR, RIDAGEYR, RIDRETH3, DMDBORN4, DMDCITZN, DMDYRSUS, DMDEDUC3, DMDEDUC2, DMDHHSIZ, INDHHIN2) %>%
  rename(gender = RIAGENDR, age = RIDAGEYR, raceth = RIDRETH3, birth_country = DMDBORN4, citizen = DMDCITZN, time_us = DMDYRSUS, edu_youth = DMDEDUC3, edu_adult = DMDEDUC2, hh_size = DMDHHSIZ, hh_income = INDHHIN2)
```


# Uploading the RAQ Data

```{r}
download.file("https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/CDQ_J.XPT", tf <- tempfile(), mode="wb")

# Create Data Frame From Temporary File
rose <- foreign::read.xport(tf)
    
# save as an R data frame
# TutorialUser: update the file path here to a directory where you want to save the data frame 
saveRDS(rose, file="C:\\NHANES\\DATA\\CDQ_J.rds")  
```
# Identifying Angina

## Grade 2 Angina

```{r}
rose <- rose %>%
  rename(q1 = CDQ001, q2 = CDQ002, q3 = CDQ003, q4 = 
           CDQ004, q5 = CDQ005, q6 = CDQ006, q9A = CDQ009A, q9B = CDQ009B, q9C = CDQ009C, q9D = CDQ009D, q9E = CDQ009E, q9F = CDQ009F, q9G = CDQ009G) %>%
  select(SEQN, q1, q2, q3, q4, q5, q6, q9D, q9E, q9F, q9G) %>%
mutate(onetosix = (ifelse(q1 == 1 &
                            q2 == 1 &
                            q3 == 1 &
                            q4 == 1 &
                            q5 == 1 &
                            q6 == 1, "yes", "no")))
rose <- rose %>%
  mutate(fourorfive = (ifelse(onetosix == "yes" &
                               q9D == 4 | 
                               onetosix == "yes" & 
                               q9E == 5, "yes", "no")))

rose <- rose %>%
  mutate(sixandseven = (ifelse(onetosix == "yes" &
                                 q9F == 6 &
                                 q9G == 7, "yes", "no")))

rose <- rose %>%
  mutate(grade2 = (ifelse(fourorfive == "yes" |
                            sixandseven == "yes", 
                          "yes", "no")))
```

## Grade 1 Angina

```{r}
rose <- rose %>%
  mutate(notthree = (ifelse(q1 == 1 &
                            q2 == 1 &
                            q3 != 1 &
                            q4 == 1 &
                            q5 == 1 &
                            q6 == 1, "yes", "no"))) %>%
  mutate(fourorfive1 = (ifelse(notthree == "yes" &
                               q9D == 4 | 
                               notthree == "yes" & 
                               q9E == 5, "yes", "no"))) %>%
  mutate(sixandseven1 = (ifelse(notthree == "yes" &
                                q9F == 6 &
                                q9G == 7, "yes", "no"))) %>%
  mutate(grade1 = (ifelse(fourorfive1 == "yes" |
                            sixandseven1 == "yes", 
                          "yes", "no")))
  
```

## Removing Extraneous Variables in the RAQ Dataset

```{r}
rose <- rose %>%
  select(SEQN, grade1, grade2)

rose[is.na(rose)] = "no"
```

# Uploading Depression Screener Data

```{r}
depression <- read_csv("depression.csv")
```

## Calculating Overall Depression Score

```{r}
depression <- depression %>%
  rename(q1 = DPQ010, q2 = DPQ020, q3 = DPQ030, q4 = 
           DPQ040, q5 = DPQ050, q6 = DPQ060, q7 = DPQ070, q8 = DPQ080, q9 = DPQ090) %>%
  select(SEQN, q1, q2, q3, q4, q5, q6, q7, q8, q9) %>%
  mutate(dscore = q1 + q2 + q3 + q4 + q5 + q6 + q7+ q8 + q9)
```

# Joining the Datasets

```{r}
# merging the depression and demographics datasets
 dreg <- 
  depression %>%
    left_join(demo, by = ("SEQN")) %>%
  select(-q1, -q2, -q3, -q4, -q5, -q6, -q7, -q8, -q9)
```

```{r}
# merging the cardiovascular (heart) health and demographics datasets

# note that this dataset will only account for people 40 years old and older

hreg <-
  rose %>%
  left_join(demo, by = ("SEQN"))
```

# Making Categorical Variables Factors

```{r}
dreg$citizen <- as.factor(dreg$citizen)
dreg$birth_country <- as.factor(dreg$birth_country)
dreg$gender <- as.factor(dreg$gender)
dreg$raceth <- as.factor(dreg$raceth)
dreg$edu_youth <- as.factor(dreg$edu_youth)
dreg$edu_adult <- as.factor(dreg$edu_adult)
dreg$hh_income <- as.factor(dreg$hh_income) 
hreg$citizen <- as.factor(hreg$citizen)
hreg$birth_country <- as.factor(hreg$birth_country)
hreg$gender <- as.factor(hreg$gender)
hreg$raceth <- as.factor(hreg$raceth)
hreg$edu_youth <- as.factor(hreg$edu_youth)
hreg$edu_adult <- as.factor(hreg$edu_adult)
hreg$hh_income <- as.factor(hreg$hh_income) 
```

# Making Numeric Variables

```{r}
hreg2 <- hreg %>% 
  mutate(grade1_num = ifelse(hreg$grade1 == "no", 0,1),
         grade2_num = ifelse(hreg$grade2 == "no", 0,1),
         anyangina = ifelse(hreg$grade1 == "yes" |
                              hreg$grade2 == "yes", 1,0))
```

# Creating Dreg Model

```{r}
# Note: did not bucket dscore because folks said it would make the analysis less robust; something is broken with time_us variable that I could not figure out so i excluded it from model
depression_model <- lm(dscore ~ hh_size + citizen + birth_country + gender + raceth + edu_adult + hh_income + age, data = dreg)


kable(tidy(depression_model, format = "markdown"), digits = 3)
```

# Creating Hreg Models

```{r}
hearthealth_model1 <- glm(grade1_num ~ hh_size + citizen + birth_country + gender + raceth + edu_adult + hh_income + age, data = hreg2, family = binomial)

kable(tidy(hearthealth_model1, format = "markdown"), digits = 3)

hearthealth_model2 <- glm(grade2_num ~ hh_size + citizen + birth_country + gender + raceth + edu_adult + hh_income + age, data = hreg2, family = binomial)

kable(tidy(hearthealth_model2, format = "markdown"), digits = 3)

anyangina_model <- glm(anyangina ~ hh_size + citizen + birth_country + gender + raceth + edu_adult + hh_income + age, data = hreg2, family = binomial)

kable(tidy(anyangina_model, format = "markdown"), digits = 3)
```